### 2015-05-03 writing/F'14 paper/join.tex
Sentence classification w/ TreeNNs

\begin{table}[tp]
  \centering  \small
  \setlength{\arraycolsep}{8pt}
  \renewcommand{\arraystretch}{1.1}
  \newcommand{\UNK}{\cdot}  
  $\begin{array}[t]{c@{ \ }|*{7}{c}|}
    %\hline
    \multicolumn{1}{c}{}
             & \nateq     & \natfor     & \natrev     & \natneg    & \natalt     & \natcov     & \multicolumn{1}{c}{\natind} \\
    \cline{2-8}
    \nateq  & \nateq &   \natfor &  \natrev &  \natneg &   \natalt &  \natcov &  \natind \\
    \natfor & \natfor &  \natfor &  \UNK &  \natalt &   \natalt &  \UNK &  \UNK \\
    \natrev & \natrev &  \UNK &  \natrev &  \natcov &   \UNK &  \natcov &  \UNK \\
    \natneg & \natneg &  \natcov &  \natalt &  \nateq &    \natrev &  \natfor &  \natind \\
    \natalt & \natalt &  \UNK &  \natalt &  \natfor &   \UNK &  \natfor &  \UNK \\
    \natcov & \natcov &  \natcov &  \UNK &  \natrev &   \natrev &  \UNK &  \UNK \\
    \natind & \natind & \UNK &  \UNK &  \natind &  \UNK &  \UNK &  \UNK \\
    \cline{2-8}
  \end{array}$
  \caption{In \S\ref{sec:join}, we assess our models' ability to learn to do inference over pairs of relations using the rules represented here, which are derived from the definitions of the relations in Table~\ref{b-table}.  As an example, given that $p_1 \natfor p_2$ and $p_2 \natneg p_3$, the entry in the $\natfor$ row and the $\natneg$ column lets us conclude that $p_1 \natalt p_3$. Cells containing a dot correspond to situations for which no valid inference can be drawn.} 
  \label{tab:jointable}
\end{table}

\section{Reasoning about semantic relations}\label{sec:join}

The simplest kinds of deduction in natural logic involve atomic statements 
using the relations in Table~\ref{b-table}. 
For instance, from the relation $p_1 \natrev p_2$ between two propositions, 
one can conclude that $p_2 \natfor p_1$ using the definitions of the relations directly. 
If one is also given the relation $p_1 \natrev p_3$, one can conclude that $p_1 \natrev p_2$, by basic set-theoretic reasoning (transitivity of $\natrev$). The
full set of sound such inferences on pairs of premise relations is depicted in
Table~\ref{tab:jointable}. Though these basic inferences do not involve compositional
sentence representations, any successful reasoning using compositional representations
will rely on the ability to perform sound inferences of this kind, so our first experiment studies how well each model can learn to perform them them in isolation.

% about the relations themselves that do not depend on the
% internal structure of the things being compared. For example, given
% that $a\sqsupset b$ and $b\sqsupset c$ one can conclude that
% $a\sqsupset c$ by the transitivity of $\sqsupset$, even without
% understanding $a$, $b$, or $c$. These seven relations support more
% than just transitivity: MacCartney and Manning's
% \cite{maccartney2009extended} join table defines 32 valid inferences
% that can be made on the basis of pairs of relations of the form $a R
% b$ and $b R' c$, including several less intuitive ones such as that if
% $a \natneg b$ and $b~|~c$ then $a \sqsupset c$.

\paragraph{Experiments}
We begin by creating a world model
on which we will base the statements in the train and test sets.
This takes the form of a small boolean structure in which terms denote
sets of entities from a small domain.  Fig.~\ref{lattice-figure}a
depicts a structure of this form with three entities ($a$, $b$, and $c$) and eight terms ($p_{1\text{-}8}$). We then generate a 
relational statement for each pair of terms in the model, as shown in Fig.~\ref{lattice-figure}b. 
We divide these statements evenly into train and test sets, and delete the test set
 examples which cannot be proven from the train examples, for which there is not enough information for even an ideal system to choose a correct label.
In each experimental run, we create a model with 80 terms over a domain of 7 elements, yielding a training set of 3200 relations and a test set of 
2960 relations.

We trained models with both the NN and NTN comparison functions on these
data sets.\footnote{Since this task relies crucially on the learning of a pair of vectors, no simpler version of our model is a viable baseline.} In both cases, the models are implemented as
described in \S\ref{methods}, but since the items being compared
are single terms rather than full tree structures, the composition
layer is not used, and the two models are not recursive. We simply present
the models with the (randomly initialized) embedding vectors for each
of two terms, ensuring that the model has no information about the terms
being compared except for the relations between them that appear in training.

@@

  \caption{Performance on the semantic relation experiments. These results and all other results on artificial data are reported as mean accuracy scores over five runs followed by mean macroaveraged F1 scores in parentheses.}
  \label{joinresultstable}
\end{table}

\paragraph{Results} 
The resuts (Table \ref{joinresultstable}) show that NTN is able to accurately encode the relations between the terms in the geometric relations between their vectors, 
and is able to then use that information to recover relations that 
are not overtly included in the training data. The NN also generalizes fairly well, 
but makes enough errors that it remains an open question whether 
it is capable of learning representations with these properties. 
We cannot rule out the possibility that different optimization techniques or
further hyperparameter tuning could lead an NN model to succeed here.

As an example from our test data, both models correctly labeled $p_1 \natfor p_3$, potentially learning from the training examples $\{p_1 \natfor p_{51},~p_3 \natrev p_{51}\}$ or $\{p_1\natfor p_{65},~p_3 \natrev p_{65} \}$. On another example involving comparably frequent relations, the NTN correctly labeled $p_6 \natrev p_{24}$, likely on the basis of the training examples $\{p_6 \natcov p_{28},~p_{28} \natneg p_{24}\}$, while the NN incorrectly assigned it $\natind$.

>>>>

\begin{table}[tp]
  \centering  \small
  \setlength{\arraycolsep}{8pt}
  \renewcommand{\arraystretch}{1.1}
  \newcommand{\UNK}{\cdot}  
  $\begin{array}[t]{c@{ \ }|*{7}{c}|}
    %\hline
    \multicolumn{1}{c}{}
             & \nateq     & \natfor     & \natrev     & \natneg    & \natalt     & \natcov     & \multicolumn{1}{c}{\natind} \\
    \cline{2-8}
    \nateq  & \nateq &   \natfor &  \natrev &  \natneg &   \natalt &  \natcov &  \natind \\
    \natfor & \natfor &  \natfor &  \UNK &  \natalt &   \natalt &  \UNK &  \UNK \\
    \natrev & \natrev &  \UNK &  \natrev &  \natcov &   \UNK &  \natcov &  \UNK \\
    \natneg & \natneg &  \natcov &  \natalt &  \nateq &    \natrev &  \natfor &  \natind \\
    \natalt & \natalt &  \UNK &  \natalt &  \natfor &   \UNK &  \natfor &  \UNK \\
    \natcov & \natcov &  \natcov &  \UNK &  \natrev &   \natrev &  \UNK &  \UNK \\
    \natind & \natind & \UNK &  \UNK &  \natind &  \UNK &  \UNK &  \UNK \\
    \cline{2-8}
  \end{array}$
  \caption{In \S\ref{sec:join}, we assess our models' ability to learn to do inference over pairs of labels using the rules represented here, which are derived from the definitions of the labels in Table~\ref{b-table}.  As an example, given that $p_1 \natfor p_2$ and $p_2 \natneg p_3$, the entry in the $\natfor$ row and the $\natneg$ column lets us conclude that $p_1 \natalt p_3$. Cells containing a dot correspond to situations for which no valid inference can be drawn.} 
  \label{tab:jointable}
\end{table}

\section{Reasoning about semantic labels}\label{sec:join}

The simplest kinds of deduction in natural logic involve atomic statements 
using the labels in Table~\ref{b-table}. 
For instance, from the label $p_1 \natrev p_2$ between two propositions, 
one can infer the label $p_2 \natfor p_1$ by applying the definitions of the labels directly. 
If one is also given the label $p_2 \natrev p_3$ one can conclude that $p_1 \natrev p_3$, by basic set-theoretic reasoning (transitivity of $\natrev$). The
full set of sound such inferences on pairs of premise labels is depicted in
Table~\ref{tab:jointable}. Though these basic inferences do not involve compositional
sentence representations, any successful reasoning using compositional representations
will rely on the ability to perform sound inferences of this kind, so our first experiment studies how well each model can learn to perform them them in isolation.

% about the labels themselves that do not depend on the
% internal structure of the things being compared. For example, given
% that $a\sqsupset b$ and $b\sqsupset c$ one can conclude that
% $a\sqsupset c$ by the transitivity of $\sqsupset$, even without
% understanding $a$, $b$, or $c$. These seven labels support more
% than just transitivity: MacCartney and Manning's
% \cite{maccartney2009extended} join table defines 32 valid inferences
% that can be made on the basis of pairs of labels of the form $a R
% b$ and $b R' c$, including several less intuitive ones such as that if
% $a \natneg b$ and $b~|~c$ then $a \sqsupset c$.

\paragraph{Experiments}
We begin by creating a world model
on which we will base the statements in the train and test sets.
This takes the form of a small Boolean structure in which terms denote
sets of entities from a small domain.  Fig.~\ref{lattice-figure}a
depicts a structure of this form with three entities ($a$, $b$, and $c$) and eight proposition terms ($p_1$--$p_8$). We then generate a 
labelal statement for each pair of terms in the model, as shown in Fig.~\ref{lattice-figure}b. 
We divide these statements evenly into train and test sets, and delete the test set
 examples which cannot be proven from the train examples, for which there is not enough information for even an ideal system to choose a correct label.
In each experimental run, we create a model with 80 terms over a domain of 7 elements, yielding a training set of 3200 examples and a test set of 
2960 examples.

We trained models with both the NN and NTN comparison functions on these
data sets.\footnote{Since this task relies crucially on the learning of a pair of vectors, no simpler version of our model is a viable baseline.} %+%
In both cases, the models are implemented as
described in \S\ref{methods}, but since the items being compared
are single terms rather than full tree structures, the composition
layer is not used, and the two models are not recursive. We simply present
the models with the (randomly initialized) embedding vectors for each
of two terms, ensuring that the model has no information about the terms
being compared except for the labels between them that appear in training.

@@

  \caption{Performance on the semantic label experiments. These results and all other results on artificial data are reported as mean accuracy scores over five runs followed by mean macroaveraged F1 scores in parentheses. The ``$\natind$ only'' entries reflect the frequency of the most frequent class.}
  \label{joinresultstable}
\end{table}

\paragraph{Results} 
The resuts (Table \ref{joinresultstable}) show that NTN is able to accurately encode the labels between the terms in the geometric labels between their vectors, 
and is able to then use that information to recover labels that 
are not overtly included in the training data. The NN also generalizes fairly well, 
but makes enough errors that it remains an open question whether 
it is capable of learning representations with these properties. 
It is not possible for us to rule out the possibility that different optimization techniques or
further hyperparameter tuning could lead an NN model to succeed here.

As an example from our test data, both models correctly labeled $p_1 \natfor p_3$, potentially learning from the training examples $\{p_1 \natfor p_{51},~p_3 \natrev p_{51}\}$ or $\{p_1\natfor p_{65},~p_3 \natrev p_{65} \}$. On another example involving comparably frequent labels, the NTN correctly labeled $p_6 \natrev p_{24}$, likely on the basis of the training examples $\{p_6 \natcov p_{28},~p_{28} \natneg p_{24}\}$, while the NN incorrectly assigned it $\natind$.

### 2015-05-03 writing/F'14 paper/quantifiers.tex
Sentence classification w/ TreeNNs

\paragraph{Experiments}
Our data consist of pairs of sentences generated
from a grammar for a simple English-like artificial language.
Each sentence contains a quantifier, a noun
which may be negated, and an intransitive verb which may be
negated. We use the quantifiers \ii{some}, \ii{most}, \ii{all},
\ii{two}, and \ii{three}, and their negations \ii{no}, \ii{not-all},
\ii{not-most}, \ii{less-than-two}, and \ii{less-than-three}, and also
include five nouns, four intransitive verbs, and the negation symbol
\ii{not}. In order to be able to define relations between sentences
with differing lexical items, we define the lexical relations for
each noun--noun pair, each verb--verb pair, and each
quantifier--quantifier pair. The grammar then generates pairs of
sentences and calculates the relations
between them.  
Some examples of these data are provided in (\ref{p1}--\ref{p3}) below.

@@

% To assign relation labels to sentence pairs, we built a small
% task-specific implemenation of MacCartney's logic that can
% accurately label sentences of this restricted language. The logic is
% not able to derive all intuitively true relations of this language,
% and fails to derive a single unique relation for certain types of
% statement, including De Morgean's laws (e.g. \ii{(all pets) growl
% $\natneg$ (some pet) (not growl)}), and we simply discard these
% examples. Exhaustively generating the valid sentences under this
% grammar and choosing those to which a relation label can be assigned
% yields 66k sentence pairs. Some examples of these data are provided
% in Table~\ref{examplesofdata}.

In each run, we partition the full set of possible sentences under the grammar into training and test sentences, and label all of the pairs from within each set to generate a training set of 27k pairs and a test set of 7k pairs. Because the model doesn't see the test sentences at training time, it cannot directly use the kind of reasoning described in \S\ref{sec:join}, each model must learn a
complete reasoning system for the limited language and logic presented
in the training data, and must be able
to recognize all of the lexical relations between the nouns, verbs,
and quantifiers and how they interact. For instance, it might see
pairs like \eqref{p1} and \eqref{p2} in training and be required to 
then label \eqref{p3}.

>>>>

\paragraph{Experiments}
Our data consist of pairs of sentences generated
from a grammar for a simple English-like artificial language.
Each sentence contains a quantifier, a noun
which may be negated, and an intransitive verb which may be
negated. We use the quantifiers \ii{some}, \ii{most}, \ii{all},
\ii{two}, and \ii{three}, and their negations \ii{no}, \ii{not-all},
\ii{not-most}, \ii{less-than-two}, and \ii{less-than-three}, and also
include five nouns, four intransitive verbs, and the negation symbol
\ii{not}. In order to be able to define labels between sentences
with differing lexical items, we define the lexical labels for
each noun--noun pair, each verb--verb pair, and each
quantifier--quantifier pair. The grammar then generates pairs of
sentences and calculates the labels
between them.  
Some examples of these data are provided in (\ref{p1}--\ref{p3}) below.

@@

% To assign label labels to sentence pairs, we built a small
% task-specific implemenation of MacCartney's logic that can
% accurately label sentences of this restricted language. The logic is
% not able to derive all intuitively true labels of this language,
% and fails to derive a single unique label for certain types of
% statement, including De Morgean's laws (e.g. \ii{(all pets) growl
% $\natneg$ (some pet) (not growl)}), and we simply discard these
% examples. Exhaustively generating the valid sentences under this
% grammar and choosing those to which a label label can be assigned
% yields 66k sentence pairs. Some examples of these data are provided
% in Table~\ref{examplesofdata}.

In each run, we partition the full set of possible sentences under the grammar into training and test sentences, and label all of the pairs from within each set to generate a training set of 27k pairs and a test set of 7k pairs. Because the model doesn't see the test sentences at training time, it cannot directly use the kind of reasoning described in \S\ref{sec:join}, each model must learn a
complete reasoning system for the limited language and logic presented
in the training data, and must be able
to recognize all of the lexical labels between the nouns, verbs,
and quantifiers and how they interact. For instance, it might see
pairs like \eqref{p1} and \eqref{p2} in training and be required to 
then label \eqref{p3}.

### 2015-05-03 writing/S'14 paper/abstract.tex
Sentence classification w/ TreeNNs

\begin{abstract}
  Supervised recursive neural network models (RNNs) for sentence
  meaning have been successful in an array of sophisticated language
  tasks, but it remains an open question whether they can learn
  compositional semantic grammars that support logical deduction.  We
  address this question directly by for the first time evaluating
  whether each of two classes of neural model --- plain RNNs and
  recursive neural tensor networks (RNTNs) --- can correctly learn
  relationships such as entailment and contradiction between pairs of
  sentences, where we have generated controlled data sets of sentences
  from a logical grammar.  Our first experiment evaluates whether
  these models can learn the basic algebra of logical relations
  involved. Our second and third experiments extend this evaluation to
  complex recursive structures and sentences involving quantification.
  We find that the plain RNN achieves only mixed results on all three
  experiments, whereas the stronger RNTN models generalizes well in
  every setting and appears capable of learning suitable
  representations for natural language logical inference.
\end{abstract}

% We address this question directly by for the first time evaluating whether each of two classes of neural model — plain RNNs and recursive neural tensor networks (RNTNs) — can correctly learn relationships such as entailment and contradiction between pairs of sentences, where we have generated controlled data sets of sentences from a logical grammar.

>>>>

\begin{abstract}
  Supervised recursive neural network models (RNNs) for sentence
  meaning have been successful in an array of sophisticated language
  tasks, but it remains an open question whether they can learn
  compositional semantic grammars that support logical deduction.  We
  address this question directly by for the first time evaluating
  whether each of two classes of neural model --- plain RNNs and
  recursive neural tensor networks (RNTNs) --- can correctly learn
  labelships such as entailment and contradiction between pairs of
  sentences, where we have generated controlled data sets of sentences
  from a logical grammar.  Our first experiment evaluates whether
  these models can learn the basic algebra of logical labels
  involved. Our second and third experiments extend this evaluation to
  complex recursive structures and sentences involving quantification.
  We find that the plain RNN achieves only mixed results on all three
  experiments, whereas the stronger RNTN model generalizes well in
  every setting and appears capable of learning suitable
  representations for natural language logical inference.
\end{abstract}

% We address this question directly by for the first time evaluating whether each of two classes of neural model — plain RNNs and recursive neural tensor networks (RNTNs) — can correctly learn labelships such as entailment and contradiction between pairs of sentences, where we have generated controlled data sets of sentences from a logical grammar.

### 2015-05-03 writing/S'14 paper/discussion.tex
Sentence classification w/ TreeNNs

This paper evaluated two recursive models on a series of three increasingly
challenging interpretive tasks involving natural language inference:
the core relational algebra of natural logic with entailment and
exclusion; recursive propositional logic structures; and statements
involving quantification and negation. The results suggest that RNTNs,
but not plain RNNs, have the capacity to meet the challenges of these
tasks with reasonably-sized training sets. These positive results are
promising for the future of learned representation models in the
applied modeling of compositional semantics.

@@

% These experiments represent one of the first attempts to reproduce any large fragment of the behavior of a complex logic within a neural network model, and the first attempt that we are aware of to address either the encoding of lexical relations or the learning of recursive operators. This presents considerable challenges in evaluating the particular models that we choose, since we cannot rely on prior results to establish that any particular amount or type of training data is sufficient to teach any model the structure of the logic. The positive results that we have found, however, are extremely promising for the future of learned representation models in the applied modeling of meaning. We have seen that recursive neural tensor networks are able to encode lexical relations accurately and encode recursive operators. We have also seen that both RNNs and RNTNs are able to handle the meanings of quantifiers in an inference setting in at least some cases.

>>>>

This paper evaluated two recursive models on a series of three increasingly
challenging interpretive tasks involving natural language inference:
the core labelal algebra of natural logic with entailment and
exclusion; recursive propositional logic structures; and statements
involving quantification and negation. The results suggest that RNTNs,
but not plain RNNs, have the capacity to meet the challenges of these
tasks with reasonably-sized training sets. These positive results are
promising for the future of learned representation models in the
applied modeling of compositional semantics.

@@

% These experiments represent one of the first attempts to reproduce any large fragment of the behavior of a complex logic within a neural network model, and the first attempt that we are aware of to address either the encoding of lexical labels or the learning of recursive operators. This presents considerable challenges in evaluating the particular models that we choose, since we cannot rely on prior results to establish that any particular amount or type of training data is sufficient to teach any model the structure of the logic. The positive results that we have found, however, are extremely promising for the future of learned representation models in the applied modeling of meaning. We have seen that recursive neural tensor networks are able to encode lexical labels accurately and encode recursive operators. We have also seen that both RNNs and RNTNs are able to handle the meanings of quantifiers in an inference setting in at least some cases.

### 2015-05-03 writing/S'14 paper/recursion.tex
Sentence classification w/ TreeNNs

\paragraph{Experiments}
As in Section~\ref{sec:join}, we test this phenomenon within the
framework of natural logic, but we now replace the unanalyzed symbols
from that experiment with complex formulae. These formulae
represent a truth-functionally complete classical propositional logic:
each atomic variable denotes either $\True$ or $\False$, and the only
operators are truth-functional ones.  Table~\ref{tab:pl} defines this
logic, and Table~\ref{tab:plexs} gives some examples of relational
statements that we can make in these terms. To compute these relations
between statements, we exhaustively enumerate the sets of assignments
of truth values to propositional variables that would satisfy each of
the statements, and then we convert the set-theoretic relation between
those assignments into one of the seven relations in
Table~\ref{b-table}. As a result, each relational statement represents
a valid theorem of the propositional logic.

\begin{table}[tp]
  \centering
  \begin{subtable}[t]{0.45\textwidth}
    \centering
    \begin{tabular}[t]{l l}
      \toprule
      Formula     & Interpretation \\
      \midrule
      $a$, $b$, $c$, $d$, $e$, $f$ & $\sem{x} \in \{\True, \False\}$ \\
      $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
      $(\varphi \pland \psi)$ & $\True$ iff $\False \notin \{\sem{\varphi}, \sem{\psi}\}$ \\
      $(\varphi \plor \psi)$  & $\True$ iff $\True \in \{\sem{\varphi}, \sem{\psi}\}$ \\
      \bottomrule
    \end{tabular}    
    \caption{Well-formed formulae. $\varphi$ and $\psi$
      range over all well-formed formulae, and $\sem{\cdot}$ is
      the interpretation function mapping formulae into $\{\True,
      \False\}$.}\label{tab:pl}
  \end{subtable}
  \quad
  \begin{subtable}[t]{0.45\textwidth}
    \centering
    \begin{tabular}[t]{r c l}
      \toprule
      $\plneg a$        & $\natneg$ & $a$ \\
      $\plneg \plneg a$ & $\nateq$  & $a$ \\
      $a$               & $\natfor$ & $(a \plor b)$ \\
      $a$               & $\natrev$ & $(a \pland b)$ \\
      %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
      $\plneg(\plneg a \pland \plneg b)$ & $\nateq$ & $(a \plor b)$ \\ 
      \bottomrule
    \end{tabular}
    \caption{Examples of statements about relations between
      well-formed formulae, defined in terms of sets of satisfying
      interpretation functions $\sem{\cdot}$.}\label{tab:plexs}
  \end{subtable}
  \caption{Propositional logic with natural logic relations.}  
  \label{prop-figure}
\end{table}

@@

For our experiments, we randomly generate a large set of  unique pairs 
of formulae and compute the relation that holds for each pair.
We discard pairs in which either statement is a tautology or
contradiction (a statement that is true of either all or no possible
assignments), for which none of the seven relation labels in
Table~\ref{b-table} can hold. The resulting set of formula pairs is
then partitioned into bins based on the number of logical operators in
the longer of the two formulae. We then randomly sample 15\% of each
bin for a held-out test set.

If we do not implement any constraint that the two statements being
compared are similar in any way, then the generated data are dominated
by statements in which the two formulae refer to largely separate
subsets of the six variables, which means that the $\natind$ relation
is almost always correct.  In an effort to balance the distribution of
relation labels without departing from the basic task of modeling
propositional logic, we disallow individual pairs of statements from
referring to more than four of the six propositional variables.

>>>>

\paragraph{Experiments}
As in Section~\ref{sec:join}, we test this phenomenon within the
framework of natural logic, but we now replace the unanalyzed symbols
from that experiment with complex formulae. These formulae
represent a truth-functionally complete classical propositional logic:
each atomic variable denotes either $\True$ or $\False$, and the only
operators are truth-functional ones.  Table~\ref{tab:pl} defines this
logic, and Table~\ref{tab:plexs} gives some examples of labelal
statements that we can make in these terms. To compute these labels
between statements, we exhaustively enumerate the sets of assignments
of truth values to propositional variables that would satisfy each of
the statements, and then we convert the set-theoretic label between
those assignments into one of the seven labels in
Table~\ref{b-table}. As a result, each labelal statement represents
a valid theorem of the propositional logic.

\begin{table}[tp]
  \centering
  \begin{subtable}[t]{0.45\textwidth}
    \centering
    \begin{tabular}[t]{l l}
      \toprule
      Formula     & Interpretation \\
      \midrule
      $a$, $b$, $c$, $d$, $e$, $f$ & $\sem{x} \in \{\True, \False\}$ \\
      $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
      $(\varphi \pland \psi)$ & $\True$ iff $\False \notin \{\sem{\varphi}, \sem{\psi}\}$ \\
      $(\varphi \plor \psi)$  & $\True$ iff $\True \in \{\sem{\varphi}, \sem{\psi}\}$ \\
      \bottomrule
    \end{tabular}    
    \caption{Well-formed formulae. $\varphi$ and $\psi$
      range over all well-formed formulae, and $\sem{\cdot}$ is
      the interpretation function mapping formulae into $\{\True,
      \False\}$.}\label{tab:pl}
  \end{subtable}
  \quad
  \begin{subtable}[t]{0.45\textwidth}
    \centering
    \begin{tabular}[t]{r c l}
      \toprule
      $\plneg a$        & $\natneg$ & $a$ \\
      $\plneg \plneg a$ & $\nateq$  & $a$ \\
      $a$               & $\natfor$ & $(a \plor b)$ \\
      $a$               & $\natrev$ & $(a \pland b)$ \\
      %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
      $\plneg(\plneg a \pland \plneg b)$ & $\nateq$ & $(a \plor b)$ \\ 
      \bottomrule
    \end{tabular}
    \caption{Examples of statements about labels between
      well-formed formulae, defined in terms of sets of satisfying
      interpretation functions $\sem{\cdot}$.}\label{tab:plexs}
  \end{subtable}
  \caption{Propositional logic with natural logic labels.}  
  \label{prop-figure}
\end{table}

@@

For our experiments, we randomly generate a large set of  unique pairs 
of formulae and compute the label that holds for each pair.
We discard pairs in which either statement is a tautology or
contradiction (a statement that is true of either all or no possible
assignments), for which none of the seven label labels in
Table~\ref{b-table} can hold. The resulting set of formula pairs is
then partitioned into bins based on the number of logical operators in
the longer of the two formulae. We then randomly sample 15\% of each
bin for a held-out test set.

If we do not implement any constraint that the two statements being
compared are similar in any way, then the generated data are dominated
by statements in which the two formulae refer to largely separate
subsets of the six variables, which means that the $\natind$ label
is almost always correct.  In an effort to balance the distribution of
label labels without departing from the basic task of modeling
propositional logic, we disallow individual pairs of statements from
referring to more than four of the six propositional variables.

### 2015-05-03 writing/S'15 misc/newmodel.tex
Sentence classification w/ TreeNNs

\begin{eqnarray}
\vec{y}^{(i,j)} & = & o^{(i, j)}_{\textsc{copy-L}} \circ \vec{y}^{(i + 1, j)} +\\
&&\nonumber o^{(i, j)}_{\textsc{copy-R}} \circ \vec{y}^{(i + 1, j + 1)} +\\
&&\nonumber o^{(i, j)}_{\textsc{compose}} \circ \vec{h}^{(i, j)}
\end{eqnarray}
%
Once all of the activations in the lattice have been computed, the top activation $\vec{y}^{(1,1)}$ can be treated as the embedding vector for the full sentence. For the sentiment classification task we study here, we pass it into a softmax classifier layer to predcit the probability of each sentiment label:
%
\begin{equation}
\vec{l} = \text{softmax}(\mathbf{M_l}\,\vec{y}^{(1, 1)} + \vec{b}_l)
\end{equation}
%
The loss for the example is then defined as the sum of three parts, a log likelihood term for the predicted sentiment class label, a sum of log likelihood terms for the correct composition location decisions (scaled by the number of ``middle'' layers in the lattice for which a composition location choice must be made), and an L2 regularization term. For sentiment relation label $\rho$, composition loctation label vector $\vec{\kappa}$, concatenated parameter vector $\theta$, and regularization strength hyperparameter $\lambda$ the loss is:
%

@@

>>>>

\begin{eqnarray}
\vec{y}^{(i,j)} & = & o^{(i, j)}_{\textsc{copy-L}} \circ \vec{y}^{(i + 1, j)} +\\
&&\nonumber o^{(i, j)}_{\textsc{copy-R}} \circ \vec{y}^{(i + 1, j + 1)} +\\
&&\nonumber o^{(i, j)}_{\textsc{compose}} \circ \vec{h}^{(i, j)}
\end{eqnarray}
%
Once all of the activations in the lattice have been computed, the top activation $\vec{y}^{(1,1)}$ can be treated as the embedding vector for the full sentence. For the sentiment classification task we study here, we pass it into a softmax classifier layer to predcit the probability of each sentiment label:
%
\begin{equation}
\vec{l} = \text{softmax}(\mathbf{M_l}\,\vec{y}^{(1, 1)} + \vec{b}_l)
\end{equation}
%
The loss for the example is then defined as the sum of three parts, a log likelihood term for the predicted sentiment class label, a sum of log likelihood terms for the correct composition location decisions (scaled by the number of ``middle'' layers in the lattice for which a composition location choice must be made), and an L2 regularization term. For sentiment label label $\rho$, composition loctation label vector $\vec{\kappa}$, concatenated parameter vector $\theta$, and regularization strength hyperparameter $\lambda$ the loss is:
%

@@

\subsection{Using pretrained word embeddings}

In order to be able to take advantage of the information contained in high-dimensional pretrained word embeddings (we use the 200d vectors distributed with Glove) without the significant expense of large intermediate node representations, we use a projection neural network layer to map the pretrained embeddings  Like in (CITE: Irsoy, Cardie) this has the effect of allowing the intermediate node representations to live in a differently structured vector space from the words. This is defined as follows, where $N$ is the row index of the bottom row (equivalent to the length of the sentence), $\mathbf{M}_t$ and $\vec{b_t}$ are the parameters of the embedding transformation layer, $\mathbf{E}$ is the matrix of pretrained embeddings, $\delta$ is a one-hot vector with a 1 at index $w$, and $w$ is the index into that matrix for the word at position $j$:

\begin{equation}
\vec{y}^{(N,j)} = \tanh(\mathbf{M_t} \mathbf{E} \delta_{w} + \vec{b_t})
\end{equation}
